---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![tic](https://github.com/tgve/tgver/actions/workflows/tic.yml/badge.svg)](https://github.com/tgve/tgver/actions/workflows/tic.yml)

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)

# tgver

This is the R package for TGVE front end npm package `tgve`. The R pacakge is developed to:
1. do interactive geospatial analysis and visuaulization fast
2. use R's echosystem to drive (1)
3. facilitate application deployment for production

## Example
This is a Markdown (GitHub) document generatd using an Rmarkdown (Rmd) document, if the Rmd is rendered to a HTML output, then using `knitr::include_url` we should see the TGVE embeded (see the live examples in the vignette).
```{r tgve, eval=FALSE}
# start a tgve instance before embedding it
tgver::tgve(background = TRUE)
knitr::include_url("http://127.0.0.1:8000")
# or use the public one
# knitr::include_url("https://tgve.github.io/eatlas-template/")
```

## TGVE interactive
Similarly, within HTML outputs we can explore (current version) `sf` objects:

```{r explore, eval=FALSE}
# using sf
nc = sf::st_read(system.file("shape/nc.shp", package="sf"))
tgver::explore(sf)
```

# Development plan
The current plan is to do use a different approach of using the TGVE npm package in this R package compared to how it is used in `geoplumber`. This is done mainly by bundling latest releases of the npm package within this package. This means not relying on npm/yarn tools for JavaScript (JS) building and bundling. As the `tgve` npm package has (see the docs) API settings to explore geospatial data, we can use R to do this without relying on the underlying JS toolchains.

We can then look at the required functions to enable workflows. For instance, a typical case would be for the package to look at a data directory and understand the data format and serve it to the front end to consume without JS builds.

Another use case would be to enable embedding `tgve` within Rmarkdowns and indeed Jupyter notebooks (both with challenges). At the moment this looks like it needs changes to be made to the npm package so that not only API and environement variables are accepted but also URL (query) variables are also accepted. If this can be achieved, then we can theoretically embed the HTML bunlded in this package within Rmarkdowns and any other browser based data science workflows.

The third use case would be to see how bundled/downloaded native apps (see the docs) generated using Rust based Tauri apps can also be driven from R or the platform (operating system) as this is not possible from browsers (of course this is possible within NodeJS).

## Funding

The project is led by Dr [Nik
Lomax](https://environment.leeds.ac.uk/geography/staff/1064/dr-nik-lomax)
and Dr [Roger
Beecham](https://environment.leeds.ac.uk/geography/staff/1003/dr-roger-beecham)
and funded by the EPSRC via the Alan Turing Institute AI for Science and
Government Programme, Grant/Award Number: EP/T001569/1.

## Contribution

Please note that the `RpackageTemplate` project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.


## Package Status

This package is part of ongoing research at the University of Leeds, it is provided "as is" and is likely to be updated and changed without warning to meet the research needs of the University.
